A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2024-01-10 paper/reviews/response.tex

\maketitle

%\tableofcontents

\thispagestyle{empty}
%\bigskip

The thoughtful comments from the two reviewers are much appreciated.  We have revised the manuscript in response. For the convenience of the Editor and Referees we include a latexdiff to show the specific changes.

@@

\begin{review}
The weak point of this work is that there is no convergence proof for the proposed algorithms. Intuitively, the constraint decomposition given in this paper is much tighter. It would give a better correction during the iterations, but this is not shown theoretically nor numerically.
\end{review}

\begin{response}
We agree that this is the weakest point of the work.  Sadly, a convergence proof is not yet available.  We believe that the algorithm is sufficiently powerful as to be worthwhile publishing, and that our numerical demonstrations adequately support this claim. We also hope that by publishing this algorithm we might stimulate theoretical investigations by others.
\end{response}

\begin{review}
The proposed FMG solver is an optimal solver as claimed by the authors. It also improves CD method of [32] in several aspects. The authors need to supply comparisons between the two methods in the numerical section.
\end{review}

\short{Our numerical demonstrations support the claim of optimality, and the reviewer seems to agree.  We already supply comparisons in Section 8; in Example 8.1 we compare to results from [22] (Graeser \& Kornhuber, 2009), who have implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, whereas the results in [32] use rather different constraint decompositions.  Our achieved convergence rate is about $0.10$, comparable to those in [22], whereas [32] achieves $>0.70$ on the same ``ball'' problem at comparable resolution.  (Smaller numbers are better here.)  On the other hand, the improvement seen in this simplest Laplacian obstacle problem is largely from our better smoother.}

@@

\begin{review}
Good if the authors can plot the error of the numerical solutions, i.e. $u^J - u^*$ and its convergence order with respect to mesh size $h_J$ with different mesh size $h_J$. This observed convergence order could be beneficial for people working with numerical analysis.
\end{review}

\short{We now report the convergence order of $\|u^\star - u^J\|$ for the two cases where the exact solution is known, namely the ``ball'' problem in Example 8.1 and the $p$-Laplacian problem in Example 8.2.}

@@

\short{We have made an effort to clarify the presentation of the FAS idea. Regarding existing literature, it remains a relatively awkward idea to explain; apparently our presentation is not exceptional.}

\begin{review}
FMG is not an iterative method, yet you start with ``FASCD \dots by a small, mesh-independent number of FMG iterations \dots''.

The FAS algorithm with V-cycles attached to make an iterative method to arbitrary solver tolerance is fine, but reporting multiple iteration is not really correct.  Right?  It is FMG + k V-cycle iterates.  Maybe report ``extra V-cycles in FASCD'', which should be 0, but you ``over solve'' to compare with others.  This is subtle and maybe could be put in the caption of Table 2 to reduce the text for the column heading from what is now ``FMG'' (``FMG V-cycles'' maybe).
\end{review}

\short{The reviewer is correct, and we have settled on a presentation close to that suggested.  As explained now in the text, notation ``$+n$'' is now used when $n$ additional V-cycles were required to satisfy the stopping criterion (7.3).  (This is documented in Section 7 before Algorithm 7.1, and at first use in Example 8.1.)  Thus ``$+0$'' appears when the FMG cycle itself sufficed, e.g.~(new) Table 3. }

@@

\begin{review}
This issue becomes a real problem in Table 5 where you devide the solver time by what you call the FASCD iterations.  This does not make sense. I would suggest, since you are just looking at time per ``iteration'', that you just do full MG and forget about iterating.
\end{review}

\short{The reviewer is correct that in Table 5 the time-divided-by-iterations column was not helpful, and we have removed it.  The ``iterations'' column is now ``$+n$'' as described above.  We believe this addresses the reviewer's very reasonable criticism.}

@@

\begin{review}
FASCD is different in that you have to do strange things with the constraints in the down smoothing, so it is not surprising that V(1,0) fails (table 3) and V(1,1) is WORSE that V(0,1) for on test, which indicates a certain fragility with the down smoothing process. It seems that should be discussed, but I don't see this.
\end{review}

\begin{response}
The description of ``strange things with the constraints in the down smoothing'' is understandable, but this is inherited from reference [22].  As we try to explain in Section 5, in down-smoothing the constraint sets must be small; this characterizes the cycles in [22].  We have improved upon [22] by recognizing that such small sets can be avoided in up-smoothing.

One point of the Example in question (i.e.~Example 8.2) is, indeed, that small constraint sets are bad for the kind of highly nonlinear VI problems we are interested in solving.  (Note that [22] does not address $p$-Laplacian cases, for other than the classical $p=2$ case, while we address the $1.5$-Laplacian and a doubly-degenerate version of the $4$-Laplacian.)  Exactly why convergence is damaged by too-small sets, in a case like this, might be revealed by a convergence proof, i.e.~the proof we don't have.

The subtle and very grid-dependent difference between the better V(0,1) and V(1,1) cycles is also something we do not understand.  It may be connected to the location of the free boundary relative to the nearest grid point, but this is too speculative for us to feel comfortable including as a remark in the manuscript.
\end{response}

### 2024-01-10 paper/fascd.tex
tiny changes to language; this is being re-submitted

Iteration results in Table \ref{tab:results:fastplap1d} show a strong contrast between $V(1,0)$ and $V(0,1)$ results, chosen by setting \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively.  The $V(1,0)$ cycles, essentially Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.  By contrast, $V(0,1)$ cycles, which up-smooth in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles.  FMG is even more efficient; a single iteration suffices at all resolutions.   Note that measured $L^\infty(\Omega)$ convergence is $O(h^{1.99})$; compare \cite{ChoeLewis1991}.
\end{example}

@@

Firedrake applications can exploit the PETSc \cite{Balayetal2023} library's parallel solvers.  The above Examples, which were run in serial, also run efficiently in parallel.  The next Examples apply parallel solvers and investigate the weak scaling of FMG.

@@

For a weak-scaling study, over $P$ processes we fixed the number of degrees of freedom per process at $m_J/P \approx 4.1 \times 10^5$ and solved with $J+1=6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer. Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in run time with increasing core counts, perhaps due to the use of single-level additive Schwarz in the Krylov iteration.  Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger in area than the current Greenland ice sheet ($1.7$ million $\text{km}^2$), at uniform $87.5$ m resolution.

>>>>

Iteration results in Table \ref{tab:results:fastplap1d} show a strong contrast between $V(1,0)$ and $V(0,1)$ results, chosen by setting \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively.  The $V(1,0)$ cycles, essentially Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.  By contrast, $V(0,1)$ cycles, which up-smooth in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles.  FMG is even more efficient; a single cycle suffices at all resolutions.   Note that measured $L^\infty(\Omega)$ convergence is $O(h^{1.99})$; compare \cite{ChoeLewis1991}.
\end{example}

@@

Firedrake applications can exploit the PETSc \cite{Balayetal2023} library's parallel solvers.  The above Examples, which were run in serial, also run efficiently in parallel.  The next two Examples apply parallel solvers and investigate the weak scaling of FMG.

@@

For a weak-scaling study, on $P$ processes we fixed the number of degrees of freedom per process at $m_J/P \approx 4.1 \times 10^5$ and solved with $J+1=6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer. Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in run time with increasing core counts, perhaps due to the use of single-level additive Schwarz in the Krylov iteration.  Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger in area than the current Greenland ice sheet ($1.7$ million $\text{km}^2$), at uniform $87.5$ m resolution.
